---
title: "Spaceship Titanic R Project"
output: html_notebook
---



```{r}
library(tidyverse)
library(Amelia)
library(GGally)
library(ggplot2)
library(reshape2)
library(readr)
```

```{r}
train <- read_csv("train.csv")
```


```{r}
head(train)

str(train)

summary(train)
```


```{r}
missing_values_df <- is.na(train)
missing_values_df <- melt(missing_values_df)
colnames(missing_values_df) <- c('row', 'variable', 'missing')

# Plot missing values
ggplot(missing_values_df, aes(x = variable, y = row)) +
  geom_raster(aes(fill = missing)) +
  scale_fill_manual(values = c('black', 'yellow')) +
  theme_minimal() +
  labs(title = 'Missing values in the training data', x = 'Variable', y = 'Row') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Save the plot
ggsave("missing_values_plot.png", width = 10, height = 8)

```

The Distribution of passengers by HomePlanet.
```{r}
homeplanet_distribution <- train %>% 
  group_by(HomePlanet) %>% 
  summarise(Count = n())

ggplot(homeplanet_distribution, aes(x = HomePlanet, y = Count)) +
  geom_bar(stat = "identity", fill = "blue") +
  theme_minimal() +
  labs(title = "Distribution of Passengers by HomePlanet", x = "HomePlanet", y = "Count")

ggsave("homeplanet_distribution.png", plot = last_plot(), width = 8, height = 6)
```

The Distribution of passengers by Age.
```{r}
ggplot(train, aes(x = Age)) +
  geom_histogram(binwidth = 5, fill = "green", color = "black") +
  theme_minimal() +
  labs(title = "Age Distribution of Passengers", x = "Age", y = "Count")

ggsave("age_distribution.png", plot = last_plot(), width = 8, height = 6)
```

Examining the relationship between Age and Transported.
```{r}
ggplot(train, aes(x = Age, fill = Transported)) +
  geom_histogram(binwidth = 10, position = "dodge") +
  theme_minimal() +
  labs(title = "Age Distribution by Transported Status", x = "Age", y = "Count")

ggsave("age_transportation_status.png", plot = last_plot(), width = 8, height = 6)
```

Examining the relationship between HomePlanet and Transported.
```{r}
ggplot(train, aes(x = HomePlanet, fill = Transported)) +
  geom_bar(position = "dodge") +
  theme_minimal() +
  labs(title = "Transported Status by HomePlanet", x = "HomePlanet", y = "Count")

ggsave("homeplanet_transportation_status.png", plot = last_plot(), width = 8, height = 6)
```

Examining the correlation between different spending categories and Transported.
```{r}
spending_categories <- c("RoomService", "FoodCourt", "ShoppingMall", "Spa", "VRDeck")

for (category in spending_categories) {
  plot <- ggplot(train, aes_string(x = category, fill = "Transported")) +
    geom_histogram(binwidth = 2000, position = "dodge") +
    theme_minimal() +
    labs(title = paste(category, "Spending by Transported Status"), x = category, y = "Count")
  
  # Save the plot
  ggsave(paste0(tolower(category), "_spending_transported.png"), plot = plot, width = 10, height = 8)
}
```


## Predicting Transported in 'test.csv'. 
Based on the analysis on ,train.csv', we have the following rules:
1. Passengers from Europa are more likely to be trasported.
2. Passengers with low spending on RoomService, ShoppingMall, Spa, and VRDeck are more likely to be transported.

```{r}
test <- read_csv("test.csv")
```

```{r}
head(test)
```

```{r}
test$Transported <- FALSE

test$Transported[test$HomePlanet == 'Europa'] <- TRUE

spending_threshold <- 50
test$Transported[rowSums(select(test, RoomService, ShoppingMall, Spa, VRDeck), na.rm = TRUE) < spending_threshold] <- TRUE

submission <- test %>% select(PassengerId, Transported)

head(submission)

write_csv(submission, "submission.csv")
```

